---
title: "P8105 HW2"
author: "Zeqi Li"
date: "2024-10-02"
output: github_document
---

```{r setup, include = FALSE}
library(haven)
library(tidyverse)
```

# Problem 1

## Load dataset

```{r load_data, message = FALSE}
subway_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                     col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) |>
  janitor::clean_names() |> 
  select(line:entry | vending | ada) |> 
  mutate(entry = case_match(entry, 
                            "YES" ~ TRUE,
                            "NO" ~ FALSE))

subway_df
```
This dataset contains the following variables that describe the NYC subway 
entrance and exit system: `r colnames(subway_df)`. When loading the dataset, 
the data type of routes 8 to 11 were converted to `chr`. After loading the 
dataset, `clean_names()` was applied to convert column names to snake case, 
`select()` was applied to retain columns desired, and `mutate()` was applied to 
convert the data type of the `entry` column from `chr` to `lgl`. The current 
dataset has a dimension of `r dim(subway_df)`. 

## Question 1
```{r q1}
subway_df |> 
  distinct(line, station_name)
```
There are 465 distinct stations.

## Question 2
```{r q2}
subway_df |> 
  filter(ada == TRUE) |> 
  distinct(line, station_name)
```
There are 84 ADA complaint stations.

## Question 3
```{r q3, eval = FALSE}
subway_df |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```
The proportion of station entrances or exits without vending allow entrance is 
0.377.

## Question 4
```{r q4_a}
subway_df |> 
  pivot_longer(route1:route11,
               names_to = "route_num",
               values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()
```
60 distinct srations serve the A train.

```{r q4_b}
subway_df |> 
  pivot_longer(route1:route11,
               names_to = "route_num",
               values_to = "route") |> 
  filter(route == "A", 
         ada == TRUE) |> 
  distinct(line, station_name)
```
Of the stations that serve the A train, 17 stations are ADA compliant.


