---
title: "P8105 HW2"
author: "Zeqi Li"
date: "2024-10-02"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(haven)
library(readxl)
```

# Problem 1

## Load dataset

```{r load_data_p1, message = FALSE}
subway_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                     col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) |>
  janitor::clean_names() |> 
  select(line:entry | vending | ada) |> 
  mutate(entry = case_match(entry, 
                            "YES" ~ TRUE,
                            "NO" ~ FALSE))

subway_df
```
This dataset contains the following variables that describe the NYC subway 
entrance and exit system: `r colnames(subway_df)`. When loading the dataset, 
the data type of routes 8 to 11 were converted to `chr`. After loading the 
dataset, `clean_names()` was applied to convert column names to snake case, 
`select()` was applied to retain columns desired, and `mutate()` was applied to 
convert the data type of the `entry` column from `chr` to `lgl`. The current 
dataset has a dimension of `r dim(subway_df)`. 

## Question 1
```{r q1}
subway_df |> 
  distinct(line, station_name)
```
There are 465 distinct stations.

## Question 2
```{r q2}
subway_df |> 
  filter(ada == TRUE) |> 
  distinct(line, station_name)
```
There are 84 ADA complaint stations.

## Question 3
```{r q3, eval = FALSE}
subway_df |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```
The proportion of station entrances or exits without vending allow entrance is 
0.377.

## Question 4
```{r q4_a}
subway_df |> 
  pivot_longer(route1:route11,
               names_to = "route_num",
               values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()
```
60 distinct srations serve the A train.

```{r q4_b}
subway_df |> 
  pivot_longer(route1:route11,
               names_to = "route_num",
               values_to = "route") |> 
  filter(route == "A", 
         ada == TRUE) |> 
  distinct(line, station_name)
```
Of the stations that serve the A train, 17 stations are ADA compliant.

# Problem 2

## Load dataset
Loading and tidying up the Mr. Trash Wheel dataset.
```{r load_data_mr_tw, message = FALSE}
mr_tw_df = read_excel("data/202409 Trash Wheel Collection Data.xlsx",
                      sheet = 1,
                      skip = 1,
                      n_max = 652) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  mutate(sports_balls = as.integer(round(sports_balls)))

mr_tw_df
```

Loading and tidying up the Professor Trash Wheel dataset.
```{r load_data_prof_tw}
prof_tw_df = read_excel("data/202409 Trash Wheel Collection Data.xlsx",
                        sheet = 2,
                        skip = 1,
                        n_max = 118) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  mutate(year = as.character(year))

prof_tw_df
```

Loading and tidying up the Gwynnda Trash Wheel dataset.
```{r load_data_g_tw}
g_tw_df = read_excel("data/202409 Trash Wheel Collection Data.xlsx",
                     sheet = 4,
                     skip = 1,
                     n_max = 262) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  mutate(year = as.character(year))

g_tw_df
```
## Combining datasets
```{r combine_data_p2}
mr_tw_df = mr_tw_df |> 
  mutate(mr_tw_df, 
         trash_wheel = "Mr. Trash Wheel") |> 
  relocate(trash_wheel)

prof_tw_df = mutate(prof_tw_df, 
                    trash_wheel = "Professor Trash Wheel") |> 
  relocate(trash_wheel)

g_tw_df = mutate(g_tw_df, 
                 trash_wheel = "Gwynnda Trash Wheel") |> 
  relocate(trash_wheel)


tw_df = bind_rows(mr_tw_df,
                  prof_tw_df,
                  g_tw_df)
tw_df
```

The combined dataset contains data of the trash collected by three trash wheel 
vessels: `r unique(pull(tw_df, 1))`. These data describe the time (ie. month, 
year, date), metrics (ie. weight in tons, volume in cubic yards), and types (ie. 
plastic bottles, polystyrene, cigarette butts, etc.) of the trash collected. 
Glass bottles data are exclusive to Mr. Trash Wheel and Professor Trash Wheel, 
and sports balls data are exclusive to Mr. Trash Wheel. 

In total, this dataset has `r dim(tw_df)[1]` observations. The total weight of 
trash collected by Professor Trash Wheel is 
`r sum(pull(filter(tw_df, trash_wheel == "Professor Trash Wheel"), weight_tons))`
 tons. The total number of cigarette butts collected by Gwynnda in June 2022 is 
`r format(sum(pull(filter(tw_df, trash_wheel == "Gwynnda Trash Wheel" & month == "June" & year == "2022"), cigarette_butts)), scientific = FALSE)`.

# Problem 3

## Load dataset
```{r load_data_p3, message = FALSE, collapse = TRUE}
baker_df = read_csv("data/gbb_datasets/bakers.csv") |> 
  janitor::clean_names()

bake_df = read_csv("data/gbb_datasets/bakes.csv",
                   na = c("N/A", "UNKNOWN", "Unknown", "")) |> 
  janitor::clean_names()

result_df = read_csv("data/gbb_datasets/results.csv",
                     skip = 2,
                     na = c("NA")) |> 
  janitor::clean_names()

baker_df
bake_df
result_df
```
## Combining datasets
First, wrangle with the `baker_name` column in `baker_df` so that we have a 
variable to work with when we join all three data frames together.
```{r baker_name}
baker_df = baker_df |> 
  separate(baker_name,
           into = c("first_name", "last_name"),
           sep = " ",
           remove = FALSE)
```
Then, join the three data frames based on series, episode, and first name.
```{r combine_data_p3}
final_df = baker_df |> 
  full_join(bake_df,
            by = join_by(first_name == baker,
                         series == series)) |> 
  full_join(result_df,
            by = join_by(first_name == baker,
                         series == series,
                         episode == episode)) |> 
  select(-(first_name:last_name)) |> 
  relocate(episode, .after = series)

final_df
```

## Export data
```{r export_data}
write_csv(final_df,
          "data/gbb_datasets/final_gbb.csv")
```

Data description

Table

## Viewership data
```{r viewrship, message = FALSE}
view_df = read_csv("data/gbb_datasets/viewers.csv") |> 
  janitor::clean_names()

view_df
```
The average viewership in Season 1 was `r mean(pull(view_df, series_1), na.rm = TRUE)`. 
The average viewership in Season 5 was `r mean(pull(view_df, series_5))`.

